---
title: Never log in to your servers again with Ansible
theme: metropolis
output:
  beamer_presentation:
    theme: metropolis
    pandoc_args: [
      "-V", "fontsize=10pt"
    ]
---
`r source("config.R")`

# Where does the name "ansible" come from?

```{r enders-game-cover, fig.height=2, fig.cap="A sci-fi novel featuring an ansible as a communication device."}
draw_image("img/enders-game-cover.jpg")
```

# Use Ansible to talk to your fleet

```{r ansible-orchestra, engine="dot"}
digraph {
  fontname=Helvetica;
  node[fontname=Helvetica shape=circle style=filled];
  control -> {web1, web2, web3, db1, db2};
  control[fillcolor="#fc8d62"];
  web1, web2, web3[fillcolor="#66c2a5"];
  db1, db2[fillcolor="#8da0cb"];
}
```

```bash
# display free disk space
control/$ ansible web -a "df"
control/$ ansible db  -a "df"
control/$ ansible all -a "df"
```

# Ad-hoc commands

```bash
# ansible [host] [-m module] [-a args]

# The default module is "command".
ansible web -a "df"
ansible web -m command -a "df"

# For piping, you need the "shell" module
ansible web -m shell -a "df | grep /dev/sda1"

# Could also use the "script" module
echo "df | grep /dev/sda1" > df_sda1.sh
ansible web -m script -a "df_sda1.sh"
```

# Inventory files

```
# contents of /etc/ansible/hosts

[web]
web1.example.com
web2.example.com
web3.example.com

[web:vars]
ansible_user=webhat

[db]
db1.example.com
db2.example.com
```

# Dynamic inventory files

Any executable file can be used as a dynamic inventory!

```
# ping ec2 servers in us-east-1d
ansible us-east-1d -i ec2.py -u ubuntu -m ping
```

# Share control of servers

```{r ansible-tower, engine="dot"}
digraph {
  fontname=Helvetica;
  node[fontname=Helvetica shape=circle style=filled];
  {player1, player2, player3} -> control -> {web1, web2, web3, db1, db2};
  player1, player2, player3[fillcolor="#e78ac3"];
  control[fillcolor="#fc8d62"];
  web1, web2, web3[fillcolor="#66c2a5"];
  db1, db2[fillcolor="#8da0cb"];
}
```

# Control a single server with Ansible

```{r ansible-simple, engine="dot", fig.width=1}
digraph {
  fontname=Helvetica;
  node[fontname=Helvetica shape=circle style=filled];
  laptop -> server;
  laptop[fillcolor="#fc8d62"];
  server[fillcolor="#66c2a5"];
}
```

# Continuous integration with Ansible

```{r ansible-ci, engine="dot", fig.width=2}
digraph {
  fontname=Helvetica;
  node[fontname=Helvetica shape=circle style=filled];
  laptop -> {VM, stage, prod};
  VM, stage, prod[size=1];
  laptop[fillcolor="#fc8d62"];
  VM[fillcolor="#8da0cb"];
  stage[fillcolor="#e78ac3"];
  prod[fillcolor="#66c2a5"];
}
```

# Provisioning on a Vagrant VM with Ansible

```ruby
# Vagrantfile
Vagrant.configure("2") do |config|
  config.vm.box = "ubuntu/trusty64"
  config.vm.provision "ansible" do |ansible|
    ansible.playbook = "vagrant.yml"
  end
end
```

# Command line tools

ansible
:   Run an ad-hoc command.

        ansible web -a "df"

playbook
:   Run a playbook.

        ansible-playbook setup.yml

vault
:   Simple encryption. Store a tough password in a password file,
    and set `ANSIBLE_VAULT_PASSWORD_FILE` environment variable.

        ansible-vault encrypt vars/secrets.yml

galaxy
:   3rd party playbooks and dependencies.

        ansible-galaxy install geerlingguy.mysql

# Ansible for DevOps

```{r ansible-for-devops}
gridExtra::grid.arrange(
  crotchet::read_image("img/ansible-for-devops-cover.jpg", width = 0.95),
  crotchet::read_image("img/jeff-geerling.jpg", width = 0.95),
  nrow = 1
)
```

# Vocabulary

playbook
:   A YAML file containing one or more plays.

play
:   A collection of tasks to run on a host or hosts.

task
:   A task is a single action with accompanying arguments,
    including a verbal description of the task.

module
:   A module is command or command wrapper.

role
:   A bundle of tasks and variables to accomplish some goal.

# Example playbook: Setting up a MySQL DB

```yaml
---
# ansible-playbook mysql.yml
- hosts: db
  become: yes
  tasks:
    - name: Install MySQL
      apt: name=mysql-server state=present
```

# YAML varieties

```yaml
- apt: name=mysql-server state=present

- apt:
    name: mysql-server
    state: present
```

# Example playbook: Setting up a MySQL DB

```yaml
---
- hosts: db
  become: yes
  vars:
    mysql_db_name: mydb
    mysql_root_password: dont-do-this
    mysql_user: pierce
    mysql_user_password: dont-do-this-either
  tasks:
    - include_role: name=geerlingguy.mysql
      vars:
        mysql_databases:
          - name: "{{ mysql_db_name }}"
        mysql_users:
          - name: "{{ mysql_user }}"
            host: "%"
            password: "{{ mysql_user_password }}"
            priv: "{{ mysql_db_name }}.*:ALL"
```

# Example playbook: Setting up a MySQL DB

```yaml
---
- hosts: db
  become: yes
  vars_files:
    - vars/main.yml
    - vars/secrets.yml
  tasks:
    - name: Configure MySQL
      include_role:
        name: geerlingguy.mysql
```

# Vars files

```yaml
---
# vars/main.yml

mysql_databases:
  - name: "{{ mysql_db_name }}"
mysql_users:
  - name: "{{ mysql_user }}"
    host: "%"
    password: "{{ mysql_user_password }}"
    priv: "{{ mysql_db_name }}.*:ALL"
```

# Typical project structure

```
root/$
  playbook.yml
  vars/
  tasks/
  roles/
```

```yaml
---
# playbook.yml
- hosts: all
  vars_files:
    - vars/main.yml
    - vars/secrets.yml
  tasks:
    - include_tasks: tasks/setup-webapp.yml
    - include_role: name=custom_role 
```

# Modules

apt
:   Package manager for Debian/Ubuntu.

        - name: Install bash and OpenSSL
          apt: name={{ item }} state=latest
          with_items:
            - bash
            - openssl

git
:   Manage git repositories.

        - git:
            repo: https://github.com/me/myrepo.git
            dest: /home/me/myrepo

pip
:   Package manager for python.

        - pip:
            requirements: /myproj/requirements.txt
            virtualenv: /venvs/myproj

# Other modules

mysql_db
:   Run commands on a MySQL db

        - name: Dump the db
          mysql_db:
            name: mydb
            state: dump
            target: /dumps/today.sql

# Other modules

fetch
:   Retrieve a file from the host(s)

        - fetch:
            src: /dumps/today.sql
            dest: /laptop/today.sql
            flat: yes

copy
:   Copy a file to the server(s)

        - copy:
            src: /laptop/yesterday.sql
            dest: /dumps/yesterday.sql

        - name: Restore the DB from a backup
          mysql_db:
            name: mydb
            state: import
            target: /dumps/yesterday.sql

